---
title: "Analysis of Marker Gene Cut-Off Efficacy"
output:
  html_document:
    df_print: paged
---

Load necessary libraries:

```{r echo=FALSE, cache=FALSE, comment=FALSE, warning=FALSE}
#load necessary libraries
library(magrittr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(limma)
library(edgeR)
library(rms)
library(useful)
library(stringr)
library(markerGeneProfile)
library(gridExtra)
library(rlist)
library(lme4)
```

Let's load the newest marker gene set information, rank them by logFC and put them into a marker gene list.
```{r}
setwd('./markerGeneEfficacy')
markers_df <-  read.csv(file = 'broad_markers_text.txt', stringsAsFactors = F)
setwd('../')


markers_df <- markers_df %>% gather(key, value, subclass)
for (subclass in unique(markers_df$value))
{
  list <- markers_df %>% filter(value == subclass)
  list <- arrange(list, -avg_log2FC)
  list <- list$gene
  subclassName <- make.names(subclass)
  print(subclassName)
  assign(subclassName, list)
  if(subclassName == make.names(unique(markers_df$value)[1])){
    final_list <- list
    print("IN")
  }
  else{
    final_list <- list(final_list, list)
  }
}
final_list <- lapply(rapply(final_list, enquote, how="unlist"), eval)
names(final_list) <- make.names(unique(markers_df$value))

setwd('./markerGeneEfficacy')
saveRDS(final_list, "broad_markers_list_logFC.rds")
setwd('../')

```
Now we've gotten the markers for each cell type ranked in order of log2FC in the format we need it to run the MGP algorithm.

We're going to get our cohort data and find the marker genes used across all cohorts for this marker list. So let's load in our cohort data. 
```{r}
#load in new cohorts QC-ed data, final count matrices and convert them to dataframes with HGNC symbols instead of ENSEMBL ids
cohorts <- c("ROSMAP", "MAYO", "MSBBM10", "MSBBM22", "MSBBM36", "MSBBM44")
for(cohort in cohorts){
  if(str_detect(cohort, "MSBB")){
    cohort <- gsub('MSB', '', cohort) 
  }
  print(cohort)
  matrix_name <- paste0(cohort, "_matrix")
  
  filename <- paste0(matrix_name, ".rds")
  setwd('./finalCountMatrices')
  matrix <- readRDS(filename)
  setwd('../')
  assign(matrix_name, matrix)

  count_df <- as.data.frame(matrix)
  count_df <- tibble:: rownames_to_column(count_df, var="Gene")
  setwd('./geneAnno')
  gene_anno <- readRDS("gene_anno.rds")
  setwd('../')
  final_df <- merge(count_df, gene_anno)
  new_gene <- final_df$Hgnc_Gene %>% make.names(unique = T) #this is ONE way of dealing with the duplicates, just making them into separate, unique names
  final_df$new_gene <- new_gene
  n_col <- ncol(count_df)
  mgp_df <- final_df[,c(n_col+2,2:n_col)]
  count_df <- mgp_df %>% rename(Gene = new_gene)
  df_name <- paste0(cohort, "_count_df")
  assign(df_name, count_df)
}
```
Let's read in the marker list we'll be using and run the MGP QC metrics to find the common marker list (the markers found in all cohorts).
```{r}
#we define our QC algorithm such that it returns a dataframe with the cell type, markers_used (list of marker genes used ' per cell type), removed_marker_ratios (list of removed marker ratios per cell type) and percent_variance_PC1 (list of variance explained by the first PC per cell type)
mgpQCMetrics <-function(count_df, mgp_markers){
  
  if(!all(c("Gene") %in% colnames(count_df))){
    stop("The count_df argument must be a df with a column named Gene (HGNC gene symbols)")
  }
  mgp_est<- markerGeneProfile::mgpEstimate(exprData=count_df,
                                                  genes=mgp_markers,
                                                  geneColName="Gene",
                                                  outlierSampleRemove=F, # should outlier samples removed. This is done using boxplot stats.
                                                  geneTransform =NULL, #function(x){homologene::mouse2human(x)$humanGene}, # this is the default option for geneTransform
                                                  groups=NULL, #if there are experimental groups provide them here. if not desired set to NULL
                                                  seekConsensus = FALSE, # ensures gene rotations are positive in both of the groups
                                                  removeMinority = TRUE)
  i= 0
  for(cell in names(mgp_markers)){
    i = i + 1
    cells_df <- mgp_est$usedMarkerExpression[i] %>% as.data.frame()
    masterlist <- paste0(rownames(cells_df), collapse=', ')
    num_markers <- length(rownames(cells_df))
    rm_marker_ratios <- mgp_est$removedMarkerRatios[i]
    if(!is.null(mgp_est$trimmedPCAs[[i]])){
      percent_variance <- ((summary(mgp_est$trimmedPCAs[[i]]))[6]) %>% as.data.frame()
      percent_variance_PC1 <- percent_variance[2,1]
    }
    else{
      percent_variance_PC1 <- NA
    }
    if(i==1){
      master_df <- data.frame( "markers_used" = masterlist, 
                        "removed_marker_ratios" = rm_marker_ratios,
                        "percent_variance_PC1" = percent_variance_PC1, 
                        "num_markers" = num_markers)  
    }
    else{
      df <- data.frame( "markers_used" = masterlist, 
                        "removed_marker_ratios" = rm_marker_ratios,
                        "percent_variance_PC1" = percent_variance_PC1, 
                        "num_markers" = num_markers)
      master_df <- rbind(master_df, df)
    }
  }
  master_df <- tibble::rownames_to_column(master_df, var = "celltype")
  return(master_df)
}


#calculate QC metrics for each marker gene list for each cohort
cohorts <- c("ROSMAP", "MAYO", "MSBBBM10", "MSBBBM22", "MSBBBM36", "MSBBBM44")
setwd('./markerGeneEfficacy')
broad_markers_list_logFC <- readRDS("broad_markers_list_logFC.rds")
marker_lists <- c("broad_markers_list_logFC")
setwd('../')

for(markers in marker_lists){
  for(cohort in cohorts){
    if(str_detect(cohort, "MSBB")){
      cohort <- gsub('MSBB', '', cohort) 
    }
    print(cohort)
    df_name <- paste0(cohort, "_count_df")
    mgpResult <- mgpQCMetrics(get(df_name), mgp_markers = get(markers))
    mgpResult$cohort <- cohort
    mgpName <- paste0("mgpQCResults",cohort, markers)
    assign(mgpName, mgpResult)
    setwd('./markerGeneEfficacy')
    saveRDS(get(mgpName), paste0(mgpName, ".rds"))
    setwd('../')
    print(mgpName)
    assign(mgpName, mgpResult)
    if(cohort =="ROSMAP"){
      all_cohorts_QC <- mgpResult
    }
    else{
      all_cohorts_QC <-rbind(all_cohorts_QC, mgpResult)
    }
  }
  all_cohorts_QC$cohort <- factor(all_cohorts_QC$cohort, levels = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO"))
  all_cohorts_QC <- arrange(all_cohorts_QC, cohort)
  setwd('./markerGeneEfficacy')
  saveRDS(all_cohorts_QC, "all_cohorts_QC.rds")
  setwd('../')
}
```

Now that we have QC metrics, which tell us which markers were used to calculate MGPs in each cohort, let's get the common markers.

```{r echo=FALSE, cache=FALSE, comment=FALSE, message=FALSE}
commonMarkerList <-function(markers, cohort_num){
    for (cell in names((markers))){
        cell_list <- all_cohorts_QC %>% filter(celltype == cell)
        start <- 1
        for(i in cohort_num){
          if(i == cohort_num[1]){
            markers1 <- unlist(strsplit(cell_list[i,2], ","))
            markers1 <- str_replace_all(string=markers1, pattern=" ", repl="")
            markers2 <- unlist(strsplit(cell_list[i,2], ","))
            markers2 <- str_replace_all(string=markers2, pattern=" ", repl="")
            common_markers <- intersect(markers1, markers2)
          }
          if(i > cohort_num[2] ){
            markersn <- unlist(strsplit(cell_list[i,2], ","))
            markersn <- str_replace_all(string=markersn, pattern=" ", repl="")
            common_markers <-intersect(common_markers, markersn)
          }
        }
        celltype_name <- eval(cell)
        celltype_name <- make.names(celltype_name)
        assign(celltype_name, common_markers)
        #only take cell types with more than 3 markers
        if(length(common_markers) > 3){
          final_common_markers <- list(final_common_markers, cell = get(celltype_name))
          names(final_common_markers) <- c('', eval(cell))
          
        }
    }
    return(final_common_markers)
  
}

marker_lists <- c("broad_markers_list_logFC")

for(markers in marker_lists){
    final_common_markers <- list()
    print(markers)
    setwd('./markerGeneEfficacy')
    all_cohorts_QC <- readRDS("all_cohorts_QC.rds")
    setwd('../')
    curr_markers <- get(markers)
    final_common_markers <- commonMarkerList(curr_markers, 1:6)
    final_common_markers <- list.flatten(final_common_markers, 
                                               use.names = TRUE, classes= "ANY")
    assign(paste0(markers, "_common_final"), final_common_markers)
    setwd('./markerGeneEfficacy')
    saveRDS(final_common_markers, paste0(markers, "_common_final.rds"))
    setwd('../')
}

```
Let's modify the QC metrics function to give us the MGPs too.
```{r}
#we define our QC algorithm such that it returns a dataframe with the cell type, markers_used (list of marker genes used ' per cell type), removed_marker_ratios (list of removed marker ratios per cell type) and percent_variance_PC1 (list of variance explained by the first PC per cell type)
mgpQCMetricsMod <-function(count_df, mgp_markers){
  
  if(!all(c("Gene") %in% colnames(count_df))){
    stop("The count_df argument must be a df with a column named Gene (HGNC gene symbols)")
  }
  mgp_est<- markerGeneProfile::mgpEstimate(exprData=count_df,
                                                  genes=mgp_markers,
                                                  geneColName="Gene",
                                                  outlierSampleRemove=F, # should outlier samples removed. This is done using boxplot stats.
                                                  geneTransform =NULL, #function(x){homologene::mouse2human(x)$humanGene}, # this is the default option for geneTransform
                                                  groups=NULL, #if there are experimental groups provide them here. if not desired set to NULL
                                                  seekConsensus = FALSE, # ensures gene rotations are positive in both of the groups
                                                  removeMinority = TRUE)
  i= 0
  for(cell in names(mgp_markers)){
    i = i + 1
    cells_df <- mgp_est$usedMarkerExpression[i] %>% as.data.frame()
    masterlist <- paste0(rownames(cells_df), collapse=', ')
    num_markers <- length(rownames(cells_df))
    rm_marker_ratios <- mgp_est$removedMarkerRatios[i]
    if(!is.null(mgp_est$trimmedPCAs[[i]])){
      percent_variance <- ((summary(mgp_est$trimmedPCAs[[i]]))[6]) %>% as.data.frame()
      percent_variance_PC1 <- percent_variance[2,1]
    }
    else{
      percent_variance_PC1 <- NA
    }
    if(i==1){
      master_df <- data.frame( "markers_used" = masterlist, 
                        "removed_marker_ratios" = rm_marker_ratios,
                        "percent_variance_PC1" = percent_variance_PC1, 
                        "num_markers" = num_markers)  
    }
    else{
      df <- data.frame( "markers_used" = masterlist, 
                        "removed_marker_ratios" = rm_marker_ratios,
                        "percent_variance_PC1" = percent_variance_PC1, 
                        "num_markers" = num_markers)
      master_df <- rbind(master_df, df)
    }
  }
  master_df <- tibble::rownames_to_column(master_df, var = "celltype")
  return(list(master_df, mgp_est))
}
```

Time to run the MGP QC metrics (modified function) on loop with mgp calc as well.

```{r}

setwd('./markerGeneEfficacy')
broad_markers_list_logFC_common_final <- readRDS("broad_markers_list_logFC_common_final.rds")
setwd('../')
marker_lists <- ("broad_markers_list_logFC_common_final")
cohorts <- c("ROSMAP", "MAYO", "MSBBBM10", "MSBBBM22", "MSBBBM36", "MSBBBM44")

for(markers in marker_lists){
  for(cohort in cohorts){
    if(str_detect(cohort, "MSBB")){
      cohort <- gsub('MSBB', '', cohort) 
    }
    print(cohort)
    df_name <- paste0(cohort, "_count_df")
    
    if(cohort == "ROSMAP"){
        covars <- c("sex","age_death")
        }
      else{
        covars <- c("msex","AgeAtDeath")
      }
      setwd('./QCpipelineResults')
      v_name <-  paste0("v_", cohort)
      voom <- readRDS(paste0(v_name, ".rds"))
      assign(v_name, voom)
      setwd('../')
      pheno_df <- voom$targets
      if(cohort == "ROSMAP" || cohort == "MAYO"){
        pheno_df<- pheno_df %>% 
          rename(
            projid = SampleID,
          )
      }
      else{
        pheno_df<- pheno_df %>% 
          rename(
            projid = sampleIdentifier,
          )
      }
    
    for(increment in seq(from=10, to=100, by=10)){
      mgp_markers <- lapply(get(markers), `[`, 1:increment)
      mgp <- mgpQCMetricsMod(get(df_name), mgp_markers = mgp_markers)
      
      estimates <- mgp[[2]]$estimates
      for(estimate in names(estimates)){
        print(estimate)
        mgp_est_cell <- as.data.frame(estimates[estimate])
        names(mgp_est_cell) <- estimate
        if(estimate == names(estimates)[1]){
          mgp_est_df <- mgp_est_cell
        }
        else{
          mgp_est_df <- cbind(mgp_est_df, mgp_est_cell)
        }
      }
      
      
      mgp_est_df <- rownames_to_column(mgp_est_df, var = "subjectID")
      mgp_est_df$cohort <- cohort
      mgp_est_df$increment <- increment

      mgp_result <- mgp[[1]]
      mgp_result$cohort <- cohort
      mgp_result$increment <- increment
      
      mgp_df <- mgp[[2]]
      mgp_df <- mgp_df$estimates %>% as.data.frame() 
      colnames(mgp_df) <- names(get(markers))
      mgp_df <- mgp_df %>% tibble::rownames_to_column(var = 'projid')
  
      # merge mgp data frame with sample metadata data frame
      mgp_df = merge(pheno_df, mgp_df, by = 'projid')
      
      for(cell in names(get(markers))){
        mgp_df[,cell] <- as.numeric(scale(mgp_df[,cell], center = TRUE, scale = TRUE))
      }
      mgp_df$increment <- increment
      mgp_name <- paste0("mgp_",cohort, "_", increment)
      mgp_ZScored_name <- paste0(mgp_name, "_ZScored")
      assign(mgp_ZScored_name, mgp_df)
      setwd('./markerGeneEfficacy')
      saveRDS(mgp_df, paste0(mgp_ZScored_name, ".rds"))
      setwd('../')

      
      if(increment == 10){
         mgp_percent_variance <- data.frame("celltype" = mgp_result$celltype,
                               "percent_variance" = mgp_result$percent_variance_PC1,
                               "cohort" = mgp_result$cohort,
                               "marker_cap" = mgp_result$increment,
                               "markers_used" = mgp_result$num_markers)
         final_est_df <- mgp_est_df
      }
      else{
        curr_percent_variance <- data.frame("celltype" = mgp_result$celltype,
                               "percent_variance" = mgp_result$percent_variance_PC1,
                               "cohort" = mgp_result$cohort,
                               "marker_cap" = mgp_result$increment,
                               "markers_used" = mgp_result$num_markers)
        mgp_percent_variance <-rbind(mgp_percent_variance, curr_percent_variance)
        final_est_df <- rbind(final_est_df, mgp_est_df)
      }
    }
    name <- paste0("mgpPercentVar",cohort, markers)
    print(name)
    assign(name, mgp_percent_variance)
    setwd('./markerGeneEfficacy')
    saveRDS(get(name), paste0(name, ".rds"))
    setwd('../')
    if(cohort =="ROSMAP"){
      all_cohort_PV <- mgp_percent_variance
      all_mgp_est <- final_est_df
    }
    else{
      all_cohort_PV <-rbind(all_cohort_PV, mgp_percent_variance)
      all_mgp_est <- rbind(all_mgp_est, final_est_df)
    }
  }
  all_cohort_PV$cohort <- factor(all_cohort_PV$cohort, levels = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO"))
  all_cohort_PV <- arrange(all_cohort_PV, cohort)
  all_mgp_est$cohort <- factor(all_mgp_est$cohort, levels = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO"))
  all_mgp_est <- arrange(all_mgp_est, cohort)
  setwd('./markerGeneEfficacy')
  saveRDS(all_cohort_PV, "all_cohort_PV.rds")
  saveRDS(all_mgp_est, "all_mgp_est.rds")
  setwd('../')
}

```
Let's plot these things because they're monstrosities and I have no idea what the trend looks like.

```{r}
setwd('./markerGeneEfficacy')
all_mgp_est<- readRDS("all_mgp_est.rds")
setwd('../')

#broad_markers_list_logFC
cell_types = names(broad_markers_list_logFC)
marker_efficacy_by_mgps <- lapply(cell_types,function(celltype) {
      print(celltype)
      mgp_plot <- 
        ggplot(all_mgp_est, aes(x=increment, y=get(celltype), color=cohort))+
        geom_point()+
        facet_grid(~cohort, scale = 'free_x', space = 'free_x') +
        labs(title= paste0(celltype, " Markers Included Vs. MGPs"),x="Top Markers Used", y = "MGP estimate")+
        theme_minimal()
      return(mgp_plot)
  })
n <- length(marker_efficacy_by_mgps)
nCol <- floor(sqrt(n))

setwd('./markerGeneEfficacy')

pdf(file = "graphed_mgps.pdf", 
    width = 25, # The width of the plot in inches
    height = 30)
print(do.call("grid.arrange", c(marker_efficacy_by_mgps, ncol=nCol)))
dev.off()
setwd('../')
```

```{r}
setwd('./markerGeneEfficacy')
all_cohort_PV <- readRDS("all_cohort_PV.rds")
setwd('../')
marker_efficacy_plot <- 
  ggplot(all_cohort_PV, 
         aes(x=markers_used, y=percent_variance, 
             group = celltype, color=celltype))+
  geom_line() +
  geom_point()+
  geom_hline(yintercept = 0.35) +
  facet_grid(~cohort, scale = 'free_x', space = 'free_x') +
  labs(title="Markers Included Vs. Percent Variance Explained",x="Top Markers Used", y = "Percent Variance Explained by PC 1")+
  theme_minimal()

setwd('./markerGeneEfficacy')
saveRDS(marker_efficacy_plot, "marker_efficacy_plot.rds")

pdf(file = "marker_efficacy.pdf", 
    width = 21, # The width of the plot in inches
    height = 14)
print(marker_efficacy_plot)
dev.off()
print(marker_efficacy_plot)
setwd('../')

```

Let's now look at our Z-scored mgps and run mega-analysis. First things first we have to organzie the Z scored mgps by the increments we ran them at and group them into dfs.

```{r}
setwd('./markerGeneEfficacy')
broad_markers_list_logFC_common_final <- readRDS("broad_markers_list_logFC_common_final.rds")
setwd('../')
marker_lists <- ("broad_markers_list_logFC_common_final")
cohorts <- c("ROSMAP", "MAYO", "BM10", "BM22", "BM36", "BM44")

setwd('./rawCohortData')
BMidsAcross <- readRDS("allMSBBIDs.rds")
setwd('../')
BMidsAcross <- BMidsAcross %>% 
  rename(
    projid = sampleIdentifier
  )

for(markers in marker_lists){
  for(increment in seq(from=10, to=100, by=10)){
    for(cohort in cohorts){
      print(cohort)
      mgp_name <- paste0("mgp_",cohort, "_", increment)
      setwd('./markerGeneEfficacy')
      mgp_ZScored_name <- paste0(mgp_name, "_ZScored")
      mgp_Z_df <- readRDS(paste0(mgp_ZScored_name, ".rds"))
      assign(mgp_ZScored_name, mgp_Z_df )
      setwd('../')
      cell_types <- names(get(markers))
      if(cohort == "ROSMAP"){
      mgp_Z_df <- mgp_Z_df %>% 
        rename(
            AgeAtDeath = age_death
          )
      }
      mgp_Z_df <- mgp_Z_df %>% select(cell_types, "projid", "msex", "LOAD", "AgeAtDeath")
      mgp_Z_df$cohort <- cohort
      if(cohort == "ROSMAP"){
        mega_mgp <- mgp_Z_df
      }
      else{
        mega_mgp <- rbind(mega_mgp, mgp_Z_df)
      }
    }
  #getting overlapping identifiers for BM cohorts
  BMidsAcross <- BMidsAcross[!duplicated(BMidsAcross),]
  MGPsBM <- mega_mgp %>% filter(str_detect(cohort, "BM"))
  allBMs <- merge(MGPsBM, BMidsAcross)
  allBMs <- allBMs[,-1]
  allBMs <- allBMs%>%select(individualIdentifier,everything())
  allBMs <- allBMs %>% 
    rename(
      projid = individualIdentifier
    )
  mega_mgp <- mega_mgp %>% filter(!str_detect(cohort, "BM"))
  mega_mgp <- rbind(mega_mgp, allBMs)
  
  #save mega_mgp
  setwd('./markerGeneEfficacy')
  saveRDS(mega_mgp, paste0("megaMGP_", markers, "_", increment, ".rds"))
  assign(paste0("megaMGP_", markers, "_", increment), mega_mgp)
  setwd('../')
    
   
  }
}

```
Okay, time to calculate the associations between LOAD and the mgps for each increment.
```{r}

setwd('./markerGeneEfficacy')
broad_markers_list_logFC_common_final <- readRDS("broad_markers_list_logFC_common_final.rds")
setwd('../')
marker_lists <- ("broad_markers_list_logFC_common_final")
cohorts <- c("ROSMAP", "MAYO", "BM10", "BM22", "BM36", "BM44")

for(markers in marker_lists){
  for(increment in seq(from=10, to=100, by=10)){
    print(markers)
    print(increment)
    
    setwd('./markerGeneEfficacy')
    mega_mgp <- readRDS(paste0("megaMGP_", markers, "_", increment, ".rds"))
    assign(paste0("megaMGP_", markers, "_", increment), mega_mgp)
    setwd('../')
    covars <- c("msex", "AgeAtDeath")
    colnames(mega_mgp) <- make.names(colnames(mega_mgp))
    cell_types <- make.names(names(get(markers)))
    pathology <- ("LOAD")
    model.data <- mega_mgp
    
    LOAD_results <- sapply(cell_types,function(celltype) {
      sapply(pathology, function(pathology) {
        print(celltype)
        form <- as.formula(paste0(celltype,"~",pathology," + ", "(1 | projid )" ,
                                  " + ", "cohort" , " + ",  paste0(covars,collapse=" + "))) 
        model <- lmer(data=model.data,form)
        return(model)
      })
    })
    
    results <- sapply(cell_types,function(celltype) {
        print(celltype)
        form <- as.formula(paste0(celltype,"~" ," + ", "(1 | projid )" ," + ", "cohort" , 
                                   " +", paste0(covars,collapse=" + "))) 
        model2 <- lmer(data=model.data,form)
        return(model2)
      })
    
    for(cell in cell_types){
      mod1Name <- paste0(cell, ".LOAD")
      mod2Name <- cell
      print(mod1Name)
      print(mod2Name)
      significance <- (anova(LOAD_results[mod1Name][[1]], results[mod2Name][[1]]))$`Pr(>Chisq)`[2]
      confInt <- confint(LOAD_results[mod1Name][[1]],level = 0.95,  oldNames=FALSE)
      upperBound <- confInt[4,2]
      lowerBound <- confInt[4,1]
      
      if(cell == cell_types[1]){
       celltype_sig <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2] ,
                                          "lowerBound" = lowerBound, "upperBound" = upperBound)
      }
      else{
        temp <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2],
                           "lowerBound" = lowerBound, "upperBound" = upperBound)
        celltype_sig <- rbind(celltype_sig, temp)
      }
    }
    
    celltype_sig$fdr <- p.adjust(celltype_sig$significance, method="fdr")
    celltype_sig$bonf <- p.adjust(celltype_sig$significance, method="bonferroni")
    celltype_sig$SIG <- celltype_sig$fdr <0.05
    celltype_sig$SIGBONF <- celltype_sig$bonf <0.05
    
    
    setwd('./markerGeneEfficacy')
    saveRDS(celltype_sig, paste0("mega_results_", markers,"_", increment, ".rds"))
    assign(paste0("mega_results_", markers, "_", increment), celltype_sig)
    setwd('../')
  }
}
```
Alright, let's plot.
```{r}
for(markers in marker_lists){
  for(increment in seq(from=10, to=100, by=10)){
    print(markers)
    
    setwd('./markerGeneEfficacy')
    mega_mgp_results <- readRDS(paste0("mega_results_", markers,"_", increment, ".rds"))
    assign(paste0("mega_mgp_results_", markers, "_", increment), mega_mgp_results)
    setwd('../')
    
    all_beta_mega = mega_mgp_results
    all_beta_mega$ub = all_beta_mega$beta + all_beta_mega$std.err
    all_beta_mega$lb = all_beta_mega$beta - all_beta_mega$std.err
    
    
    #add the *** label for significant vs. not significant
    annotation_label_mega <- all_beta_mega
    annotation_label_mega$mark <- ifelse(annotation_label_mega$bonf <0.05,"***", "")
    
    
    mega_analysis_plot = all_beta_mega %>% 
    ggplot(aes(x = celltype, y = beta)) + 
    geom_hline(yintercept = 0) + 
    geom_bar(stat = "identity", show.legend = FALSE) + 
    scale_fill_manual() + 
    geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
    ylab('LOAD (Beta)') + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    ggtitle(paste0("Mega Analysis Results for ", increment, " Markers Marker List")) +
    geom_text(x = annotation_label_mega$celltype,  y = 0.3, 
              label = annotation_label_mega$mark, 
              colour = "black", size=6)

    print(mega_analysis_plot)
    setwd('./markerGeneEfficacy')
    ggsave(paste0("mega_analysis_", markers, "_", increment, "_plot", ".png"))
    setwd('../')
  }
}
```
Let's look at the marker gene associations with LOAD.

```{r}
frenchFryPlot<-function(AD_coef, AD_pval, marker_list, cohort){
  
  
  pathology_df <- AD_pval
  
  marker_df <-  data.frame(unlist(marker_list, use.names=F),rep(names(marker_list),
                                                                lengths(marker_list)))
  colnames(marker_df) <- c("marker", "celltype")
  for (marker in names(marker_list)){
    df <- pathology_df %>% filter(marker == marker)
    if (marker == names(marker_list)[1]){
      result <- df
    }
    else{
      result <- rbind(result, df)
    }
  }
  
  final_result <- merge(result, AD_coef, by="gene")
  final_result$signedP <- -log10(final_result$pval) *(sign(final_result$coef))
  
  merge_markers<- marker_df %>% rename( gene = marker)
  final_result <- merge(merge_markers, final_result)
  final_result$cohort <- c(cohort)
  return(unique(final_result))
  
}

setwd('./markerGeneEfficacy')
broad_markers_list_logFC_common_final <- readRDS("broad_markers_list_logFC_common_final.rds")
setwd('../')
marker_lists <- ("broad_markers_list_logFC_common_final")
cohorts <- c("ROSMAP", "MAYO", "BM10", "BM22", "BM36", "BM44")

setwd('./geneAnno')
gene_anno <- readRDS("gene_anno.rds")
setwd('../')

for (markers in marker_lists){
  for(cohort in cohorts){
    if(str_detect(markers, "subclass_both")){
      #if subclass_both specified run CgG markers on prefrontal and MTG on temporal
      if(cohort == "ROSMAP" | cohort =="BM10" | cohort =="BM44"){
       markers_for_plot <- subclass_both_CgG_common_final
      }
      else{
        markers_for_plot <- subclass_both_MTG_common_final
        }
    }
    else{
      markers_for_plot <- get(markers)
    }
    print(cohort)
    lmod_name <-  paste0("lmod_" ,cohort)
    
    filename <- paste0(lmod_name, ".rds")
    setwd('./cohortQCMods')
    lmod <- readRDS(filename)
    setwd('../')
    assign(lmod_name, lmod)
    eb <- eBayes(lmod,robust = T)
    
    if(cohort == "ROSMAP"){
      n=11
    }
    if(cohort == "MAYO"){
      n=24
    }
    if(cohort == "BM10"){
      n=18
    }
    if(cohort == "BM22"){
      n=21
    }
    else if(cohort == "BM36" || cohort == "BM44"){
      n=19
    }

    print(n)
    gene_v_AD  <-lmod$coefficients[,c(n)]
    gene_v_AD <- as.data.frame(gene_v_AD)
    gene_v_AD <- tibble:: rownames_to_column(gene_v_AD, var="Gene")
    
    final_df <- merge(gene_v_AD, gene_anno)
    final_df$new_gene<- final_df$Hgnc_Gene %>% make.names(unique = T) #this is ONE way of dealing with the duplicates, just making them into separate, unique names
    #add new_gene column w/ Hgnc_Gene values filtered to have no duplicates
    AD_coef <- final_df[,c(2,4)]
    colnames(AD_coef) <- c("coef", "gene")
    
    coef_df <- AD_coef #dataframe with coefficient column and HGNC gene name column 

    #get p values
    p_geneAD <- eb$p.value
    p_geneAD  <-p_geneAD[,c(n)]
    p_geneAD <- as.data.frame(p_geneAD)
    p_geneAD <- tibble:: rownames_to_column(p_geneAD, var="Gene")
    final_df <- merge(p_geneAD, gene_anno)
    new_gene <- final_df$Hgnc_Gene %>% make.names(unique = T) #this is ONE way of dealing with the duplicates, just making them into separate, unique names
    #add new_gene column w/ Hgnc_Gene values filtered to have no duplicates
    final_df$new_gene <- new_gene
    AD_pval <- final_df[,c(2,4)]
    colnames(AD_pval) <- c("pval", "gene")
    
    gene_pathology_association <- frenchFryPlot(AD_coef, AD_pval, cohort, marker_list =markers_for_plot)
    assign(paste0("fry_plot", cohort), gene_pathology_association)
  }
  fry_df <- rbind(fry_plotROSMAP, fry_plotMAYO, fry_plotBM10,
                  fry_plotBM22, fry_plotBM36, fry_plotBM44)
  fry_df$fdr <- p.adjust(fry_df$pval, method="fdr")
  fry_df$signedFDR <- -log10(fry_df$fdr) *(sign(fry_df$signedP))
  fry_df$sig <- ifelse(fry_df$pval < 0.05, "#94C973", "#808080")
  
  for (cell in names((markers_for_plot))){
    celltype_fry_df <- fry_df %>% filter(celltype == cell)
    celltype_heat_map <- ggplot(celltype_fry_df, aes(cohort, gene, fill= signedP))+
      theme_minimal() + geom_tile() + 
      scale_fill_gradient2(low="darkblue", high="darkgreen", guide="colorbar") +   
      ggtitle(paste0("Significance of \n" , cell, " ", markers , 
                     "\n marker genes per cohort")) 
    
    celltype_french <- ggplot(celltype_fry_df, aes(x= gene,y= signedP))+
      theme_minimal() + theme(axis.text.x = element_text(angle = 45)) +
      geom_bar(stat="identity", fill = celltype_fry_df$sig)+ 
      facet_wrap(~cohort, scales = 'free_x',nrow=6) + 
      ggtitle(paste0("Significance of \n" , cell, " ", markers , 
                     "\n marker genes per cohort"))
    
    celltype_heat_map
    celltype_french
    
    setwd('./frenchHeat')
    heat_name <- paste0("heat_map",make.names(cell))
    assign(heat_name, celltype_heat_map)
    print(get(heat_name))
    saveRDS(get(heat_name), paste0(heat_name, ".rds"))
    ggsave(paste0(heat_name, markers, "_plot", ".png"), width = 15, height = 12)
    
    french_name <- paste0("french_fry",make.names(cell))
    assign(french_name, celltype_french)
    print(get(french_name))
    saveRDS(get(french_name), paste0(french_name, ".rds"))
    ggsave(paste0(french_name, markers, "_plot", ".png"), width = 15, height = 20)
    setwd('../')
    
  }
}
```
