---
title: "Mega-Analysis -- FINAL"
output:
  html_document:
    df_print: paged
---
```{r echo=FALSE, cache=FALSE, comment=FALSE, warning=FALSE}
#load necessary libraries
library(magrittr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(limma)
library(edgeR)
library(rms)
library(useful)
library(stringr)
library(markerGeneProfile)
library(rlist)
library(lme4)
library(ggthemes)
library(ggsci)
library(ggpubr)
library(ggbeeswarm)
library(cowplot)
```
Now that we've loaded in the libraries lets get the marker lists we need.
```{r}
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")
for(marker in marker_lists){
  print(marker)
  if(marker != "lake" && marker != "darmanis"){
     folder <- './commonMarkerLists/'
  }
  else{
     folder <- './markerLists/'
  }
  marker_info <-  readRDS(paste0(folder, marker, '.rds'))
  assign(marker, marker_info)
}
```
We're going to create a dataframe for mega-analysis for each of the marker gene lists, which will contain all the relative cell-type proportion estimates (rCTPs) calculated for each subject in each of the cohorts with the given marker gene lists, as well as their sample identifier (projid), their age at death, their sex and their alzheimer's diagnosis. 
We've been treating each of the brain regions thus far as separate cohorts, i.e. that there is a ROSMAP, MAYO, BM10, BM22, BM36 and BM44 cohort, but the truth is the Mount Sinai cohort contains the BM10, BM22, BM36 and BM44 cohorts. This means there is overlap in subjects between these four "cohorts", i.e. that a subject *X* may have bulk-tissue RNA-seq data sampled from BM10 and BM22, and therefore these readings are not independent, as they're confounded by being from the same subject, *X*.
As of now we've been using unique identifiers in the Mount Sinai cohorts for each brain region, but we want to rename the identifiers so we can be aware of which subjects have multiple regions sampled. This way we can account for the repetition of subejcts in our mega-analysis. We're going to convert the IDs to no longer be unique across all brain regions, but to allow for us to perceive subject re-sampling when we create our mega-analysis dataframes.
```{r}
cohorts <- c("ROSMAP", "MAYO", "BM10", "BM22", "BM36", "BM44")
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")

BMidsAcross <- readRDS("./rawCohortData/allMSBBIDs.rds")
BMidsAcross <- BMidsAcross %>% 
  dplyr::rename(
    projid = sampleIdentifier
  )
    
    
for(markers in marker_lists){
  print(markers)
  folder_name <- gsub('_common_final', '', markers)
  for(cohort in cohorts){
    print(cohort)
    mgp_name <- paste0("mgp_",cohort)
    mgp_ZScored_name <- paste0(mgp_name, "_ZScored")
    mgp_Z_df <- readRDS(paste0('./mgpResults_',folder_name, '/', mgp_ZScored_name, ".rds"))
    assign(mgp_ZScored_name, mgp_Z_df )
    cell_types <- names(get(markers))
    if(cohort == "ROSMAP"){
      mgp_Z_df <- mgp_Z_df %>% dplyr::mutate(Diagnosis = 
                                               case_when( (braaksc >= 4 & ceradsc <= 2 & cogdx == 4) 
                                                          ~ 'AD',
                                                          (braaksc <= 3 & ceradsc >= 3 & cogdx == 1) 
                                                          ~ 'C',
                                                      TRUE ~ 'OTHER'))
      mgp_Z_df <- mgp_Z_df %>% dplyr::filter(Diagnosis != "OTHER")
      mgp_Z_df$LOAD <- factor(ifelse(mgp_Z_df$Diagnosis=="AD",1,0))
      mgp_Z_df <- mgp_Z_df %>% 
        dplyr::rename(
            AgeAtDeath = age_death
          )
    }
    mgp_Z_df <- mgp_Z_df %>% dplyr::select(cell_types, "projid", "msex", "LOAD", "AgeAtDeath")
    mgp_Z_df$cohort <- cohort
    if(cohort == "ROSMAP"){
      mega_mgp <- mgp_Z_df
    }
    else{
      mega_mgp <- rbind(mega_mgp, mgp_Z_df)
    }
  }
  #getting overlapping identifiers for BM cohorts
  BMidsAcross <- BMidsAcross[!duplicated(BMidsAcross),]
  MGPsBM <- mega_mgp %>% filter(str_detect(cohort, "BM"))
  allBMs <- merge(MGPsBM, BMidsAcross)
  allBMs <- allBMs[,-1]
  allBMs <- allBMs%>% dplyr:: select(individualIdentifier,everything())
  allBMs <- allBMs %>% 
    dplyr::rename(
      projid = individualIdentifier
    )
  mega_mgp <- mega_mgp %>% dplyr::filter(!str_detect(cohort, "BM"))
  mega_mgp <- rbind(mega_mgp, allBMs)
  
  #save mega_mgp
  saveRDS(mega_mgp, paste0('./mgpResults_',folder_name, '/', "megaMGP_", markers, ".rds"))
  assign(paste0("megaMGP_", markers), mega_mgp)
}
```
Now we have all the dataframes we need for the mega-analysis. Let's do it.
It's important to note that the "In8" celltype describe in the "lake" marker gene set is defined using only one gene "NMU", which is not enough for the MGP method to calculate a cell-type. Thus all it's values are N/A. We will therefore exclude it in the mega-analysis as the "In8" cell-type cannot be determined.
```{r}
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")

    
for(markers in marker_lists){
  print(markers)
  folder_name <- gsub('_common_final', '', markers)
 
  mega_mgp <- readRDS(paste0('./mgpResults_',folder_name, "/megaMGP_", markers, ".rds"))
  assign(paste0("megaMGP_", markers), mega_mgp)
  covars <- c("msex", "AgeAtDeath")
  colnames(mega_mgp) <- make.names(colnames(mega_mgp))
  cell_types <- make.names(names(get(markers)))
  if(markers == "lake"){
    #remove In8 as it is all N/A
    cell_types <- cell_types[-16]
  }
  pathology <- ("LOAD")
  model.data <- mega_mgp
  
  LOAD_results <- sapply(cell_types,function(celltype) {
    sapply(pathology, function(pathology) {
      print(celltype)
      form <- as.formula(paste0(celltype,"~",pathology," + ", "(1 | projid )" ,
                                  " + ", "cohort" , " + ",  paste0(covars,collapse=" + "))) 
      model <- lmer(data=model.data,form)
      return(model)
    })
  })
  
  results <- sapply(cell_types,function(celltype) {
      print(celltype)
      form <- as.formula(paste0(celltype,"~" ," + ", "(1 | projid )" ," + ", "cohort" , 
                                   " +", paste0(covars,collapse=" + "))) 
      model2 <- lmer(data=model.data,form)
      return(model2)
    })
  
  for(cell in cell_types){
    mod1Name <- paste0(cell, ".LOAD")
    mod2Name <- cell
    print(mod1Name)
    print(mod2Name)
    significance <- (anova(LOAD_results[mod1Name][[1]], results[mod2Name][[1]]))$`Pr(>Chisq)`[2]
    confInt <- confint(LOAD_results[mod1Name][[1]],level = 0.95,  oldNames=FALSE)
    upperBound <- confInt[4,2]
    lowerBound <- confInt[4,1]
    
    if(cell == cell_types[1]){
     celltype_sig <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2] ,
                                        "lowerBound" = lowerBound, "upperBound" = upperBound)
    }
    else{
      temp <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2],
                         "lowerBound" = lowerBound, "upperBound" = upperBound)
      celltype_sig <- rbind(celltype_sig, temp)
    }
  }
  
  celltype_sig$fdr <- p.adjust(celltype_sig$significance, method="fdr")
  celltype_sig$bonf <- p.adjust(celltype_sig$significance, method="bonferroni")
  celltype_sig$SIG <- celltype_sig$fdr <0.05
  celltype_sig$SIGBONF <- celltype_sig$bonf <0.05
  
  
  saveRDS(celltype_sig, paste0("./megaResults/mega_results_", markers, ".rds"))
  assign(paste0("mega_results_", markers), celltype_sig)
  
  
  mega_mgp_res <- mega_mgp
  for(celltype in cell_types){
    #(1 | big_brain_region / cohort)
    form <- as.formula(paste0(celltype,"~", "(1 | projid )" ," + ", "cohort" , " + ",  paste0(covars,collapse=" + "))) 
    model <- lmer(data=model.data,form)
    cell_type_residual <- data.frame(resid(model))
    mega_mgp_res[paste0(celltype, "LOADResid")] <- cell_type_residual
  }
  
  #save the residuals 
  assign(paste0("mega_mgp_res", folder_name),mega_mgp_res)
  saveRDS(mega_mgp_res, paste0('./mgpResults_',folder_name, "/mega_mgp_res.rds"))
  write.csv(mega_mgp_res,paste0('./mgpResults_',folder_name, "/mega_mgp_res.csv"))
}
  
```
We've calculated the significance of the association between each of the cell-types and the AD diagnosis variable (LOAD) in a mega-analyis. We can now plot the results.
```{r}
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")


subclass_meta <- read.csv('./subclassMeta/subclass_meta.txt')

    
for(markers in marker_lists){
  print(markers)
  folder_name <- gsub('_common_final', '', markers)
  
  mega_mgp_results <- readRDS(paste0("./megaResults/mega_results_", markers, ".rds"))
  assign(paste0("mega_results_", markers), mega_mgp_results)
  
  
  
  
  if(str_detect(markers, "subclass")){
      mega_mgp_results <- mega_mgp_results %>% dplyr::rename(subclass = celltype)
      subclass_meta$AIBS_subclass_label <- make.names(subclass_meta$AIBS_subclass_label)
      all_beta_mega = merge(mega_mgp_results,
                      subclass_meta, by.x = 'subclass', by.y = 'AIBS_subclass_label')
      all_beta_mega$class <- as.factor(all_beta_mega$AIBS_class_label)
      all_beta_mega$class <- factor(all_beta_mega$AIBS_class_label, 
                                levels = c("GABAergic", "Glutamatergic", 
                                                                "Non-neuronal"))
      all_beta_mega <- dplyr::arrange(all_beta_mega, class)
  }
  else{
    all_beta_mega = mega_mgp_results
  }
  all_beta_mega$ub = all_beta_mega$beta + all_beta_mega$std.err
  all_beta_mega$lb = all_beta_mega$beta - all_beta_mega$std.err
  
  
  
  
  
  #add the *** label for significant vs. not significant
  annotation_label_mega <- all_beta_mega
  annotation_label_mega$mark <- ifelse(annotation_label_mega$bonf <0.05,"***", "")
  
  
  
  my_colours = c('blue', 'grey', 'red', 'green', 'yellow', 'purple') 
  if(str_detect(markers, "subclass")){
    mega_analysis_plot = all_beta_mega %>% 
    ggplot(aes(x = subclass, y = beta,fill = AIBS_class_color)) + 
    geom_hline(yintercept = 0) + 
    geom_bar(stat = "identity", show.legend = FALSE) + 
    scale_fill_manual(values = my_colours) + 
    facet_grid(~AIBS_class_label, scale = 'free_x', space = 'free_x') +
    geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
    ylab('LOAD (Beta)') + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
          strip.text.x = element_text(size = 15))+
    ggtitle(paste0("Mega Analysis Results for ", folder_name, " Marker List")) +
    geom_text(x = annotation_label_mega$subclass,  y = 0.3, 
              label = annotation_label_mega$mark, 
              colour = "black", size=6)
   
  }
  else{
    mega_analysis_plot = all_beta_mega %>% 
    ggplot(aes(x = celltype, y = beta)) + 
    geom_hline(yintercept = 0) + 
    geom_bar(stat = "identity", show.legend = FALSE) + 
    scale_fill_manual(values = my_colours) + 
    geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
    ylab('LOAD (Beta)') + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    ggtitle(paste0("Mega Analysis Results for ", folder_name, " Marker List")) +
    geom_text(x = annotation_label_mega$celltype,  y = 0.3, 
              label = annotation_label_mega$mark, 
              colour = "black", size=6)
  }
  print(mega_analysis_plot)
  ggsave(paste0("./megaResults/mega_analysis_", markers, "_plot", ".png"))
}
```
Let's create case/control box plots for SST and IT rCTP residuals so we can see if there's a difference in cell type proportion changes across cohorts/brain regions.
```{r}
cohorts <- c("ROSMAP", "MAYO", "BM10", "BM22", "BM36", "BM44")
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")

    
for(markers in marker_lists){
  print(markers)
  folder_name <- gsub('_common_final', '', markers)
  full_res_indiv <- data.frame()
  full_sig_indiv <- data.frame()
  for(cohort in cohorts){
    print(cohort)
    mgp_name <- paste0("mgp_",cohort)
    mgp_ZScored_name <- paste0(mgp_name, "_ZScored")
    mgp_Z_df <- readRDS(paste0('./mgpResults_',folder_name, '/', mgp_ZScored_name, ".rds"))
    assign(mgp_ZScored_name, mgp_Z_df )
    cell_types <- make.names(names(get(markers)))
    if(markers == "lake"){
      #remove In8 as it is all N/A
      cell_types <- cell_types[-16]
    }
    if(cohort == "ROSMAP"){
      mgp_Z_df <- mgp_Z_df %>% dplyr::mutate(Diagnosis = 
                                               case_when( (braaksc >= 4 & ceradsc <= 2 & cogdx == 4) 
                                                          ~ 'AD',
                                                          (braaksc <= 3 & ceradsc >= 3 & cogdx == 1) 
                                                          ~ 'C',
                                                      TRUE ~ 'OTHER'))
      mgp_Z_df <- mgp_Z_df %>% dplyr::filter(Diagnosis != "OTHER")
      mgp_Z_df$LOAD <- factor(ifelse(mgp_Z_df$Diagnosis=="AD",1,0))
      mgp_Z_df <- mgp_Z_df %>% 
        dplyr::rename(
            AgeAtDeath = age_death
          )
    }
    colnames(mgp_Z_df) <- make.names(colnames(mgp_Z_df))
    mgp_Z_df$cohort <- cohort
    pathology <- "LOAD"
    covars <- c("msex", "AgeAtDeath")
    model.data <- mgp_Z_df
    i=0
    for(celltype in cell_types){
      form <- as.formula(paste0(celltype,"~",paste0(covars,collapse=" + "))) 
      model <- lm(data=model.data,form)
      celltype_residual <- data.frame(resid(model))
      
      model.data[paste0(celltype, "LOADResid")] <- celltype_residual
      model.data$cohort <- cohort
      
      form <- as.formula(paste0(celltype,"~",pathology," + ",paste0(covars,collapse=" + "))) 
      model <- lm(data=model.data,form)
      loadp <- (coef(summary(model))[2,4])
      
      print(celltype, loadp)
      
      if(i==0){
           sig_results <- data.frame("celltype"= celltype, "cohort" = cohort, "significance" = loadp)
      }
      else{
        sig_results_curr <- data.frame("celltype"= celltype, "cohort" = cohort, "significance" = loadp)
        sig_results <- rbind(sig_results, sig_results_curr)
      }
      i= i+1
    }
    cell_type_resids <- paste(cell_types, "LOADResid", sep="")
      full_res_indiv <- rbind(full_res_indiv, model.data[, c(cell_type_resids, "projid", "cohort", "LOAD")])
    
    full_sig_indiv <- rbind(full_sig_indiv, sig_results)
    }
    
  }
  
  
  full_res_indiv$Diagnosis <- (ifelse(full_res_indiv$LOAD == 1, "AD","C"))
  full_res_indiv$Diagnosis <- factor(full_res_indiv$Diagnosis, levels = c("C", "AD"))
  
  saveRDS(full_res_indiv, paste0('./mgpResults_',folder_name, "/full_res_", markers, ".rds"))
  assign(paste0("full_res_", markers), full_res_indiv)
  saveRDS(full_sig_indiv, paste0('./mgpResults_',folder_name, "/full_sig_", markers, ".rds"))
  assign(paste0("full_sig_", markers), full_sig_indiv)

```
Let's plot the case control boxplots now that we've calculated the residuals.
```{r}
marker_lists <- c("subclass_CgG_MTG_common_final", "lake", "darmanis", "subclass_CgG_MTG_noSST_common_final")
for(markers in marker_lists){
  print(markers)
  folder_name <- gsub('_common_final', '', markers)
  full_res <- readRDS(paste0('./mgpResults_',folder_name, "/full_res_", markers, ".rds"))
  assign(paste0("full_res_", markers), full_res)
  full_sig <- readRDS(paste0('./mgpResults_',folder_name, "/full_sig_", markers, ".rds"))
  assign(paste0("full_sig_", markers), full_sig)
  
  full_res$cohort <- factor(full_res$cohort, levels = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO"))
  full_res <- arrange(full_res, cohort)
  
  full_sig$cohort <- factor(full_sig$cohort, levels = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO"))
  full_sig <- arrange(full_sig, cohort)
  
  #add the *** label for significant vs. not significant
  annotation_label <- full_sig
  annotation_label$bonf <- p.adjust(annotation_label$significance, method="bonferroni")
  annotation_label$fdr <- p.adjust(annotation_label$significance, method="fdr")
  annotation_label$mark <- ifelse(annotation_label$significance <0.05,"*", "ns")
  
  cell_types <- make.names(names(get(markers)))
  if(markers == "lake"){
    #remove In8 as it is all N/A
    cell_types <- cell_types[-16]
  }
  
  for(cell in cell_types){
    print(cell)
    label_mark <- subset(annotation_label, celltype == cell)
    data_text <- data.frame(
      label = label_mark$mark,
      cohort   = c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO")
    )
    boxplots <- full_res %>% 
      ggplot(aes(x = Diagnosis, y = get(paste0(cell, "LOADResid")) )) + theme_minimal()+
      #geom_violin(alpha=0.4) + 
      geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
      stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
      scale_fill_brewer(palette="Set1")+
      scale_color_brewer(palette="Set1")+
      ggtitle(paste0(cell, ' rCTP residuals association with LOAD for ', markers))+
      facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
      ylab(paste0(cell, ' rCTP residuals')) + 
      xlab('')  + geom_text(
      data    = data_text,
      mapping = aes(x = 1, y = 3.5, label = label),
      hjust   = -0.1,
      vjust   = -1
    )
    assign(paste0(cell, "_boxplot_", markers), boxplots)
    print(boxplots)
    folder_name <- gsub('_common_final', '', markers)
    saveRDS(get(paste0(cell, "_boxplot_", markers)), 
            paste0('./mgpResults_',folder_name, "/case_control_", cell, "_", markers, "_boxplot.rds"), compress=TRUE)
 
    ggsave(paste0("./caseControlPlots/case_control_", cell, "_", markers, "_boxplot", ".png"))
  }
  
  
  
}
```

```{r}

subclass_meta <- read.csv('./subclassMeta/subclass_meta.txt')

marker_list <- c("subclass_CgG_MTG_common_final", "subclass_CgG_MTG_noSST_common_final")
for(markers in marker_list){
  folder_name <- gsub('_common_final', '', markers)
  
  full_res <- readRDS(paste0('./mgpResults_',folder_name, "/full_res_", markers, ".rds"))
  assign(paste0("full_res_", "markers"), full_res)
  full_sig <- readRDS(paste0('./mgpResults_',folder_name, "/full_sig_", markers, ".rds"))
  assign(paste0("full_sig_", markers), full_sig)
  
  cohort_name <- c("ROSMAP", "BM10", "BM44", "BM22", "BM36", "MAYO")
  full_res$cohort <- factor(full_res$cohort, levels = cohort_name)
  full_sig$cohort <- factor(full_sig$cohort, levels = cohort_name)
    
  full_res <- arrange(full_res, cohort)
  full_sig <- arrange(full_sig, cohort)
  #add the *** label for significant vs. not significant
  annotation_label <- full_sig
  annotation_label$bonf <- p.adjust(annotation_label$significance, method="bonferroni")
  annotation_label$mark <- ifelse(annotation_label$significance <0.05,"*", "ns")
  
  
  #get mega analysis plot
  mega_mgp_results <- readRDS(paste0("./megaResults/mega_results_", markers, ".rds"))
  assign(paste0("mega_results_", markers), mega_mgp_results)
  
  
  mega_mgp_results <- mega_mgp_results %>% dplyr::rename(subclass = celltype)
  subclass_meta$AIBS_subclass_label <- make.names(subclass_meta$AIBS_subclass_label)
  all_beta_mega = merge(mega_mgp_results,
                  subclass_meta, by.x = 'subclass', by.y = 'AIBS_subclass_label')
  all_beta_mega$class <- as.factor(all_beta_mega$AIBS_class_label)
  all_beta_mega$class <- factor(all_beta_mega$AIBS_class_label, 
                            levels = c("GABAergic", "Glutamatergic", 
                                                            "Non-neuronal"))

  
  all_beta_mega$subclass <- gsub(".", " ", all_beta_mega$subclass, fixed=TRUE)
  all_beta_mega$subclass <- factor(all_beta_mega$subclass, 
                            levels = c("LAMP5", "PAX6", "VIP", "SST", "PVALB", "IT", 
                            "L4 IT", "L5 6 NP", "L5 ET", "L6 CT", "L5 6 IT Car3", "L6b",
                            "Astrocyte", "Endothelial", "Microglia", "Oligodendrocyte",
                            "OPC", "Pericyte", "VLMC"))
    
  all_beta_mega <- dplyr:: arrange(all_beta_mega, class)
  all_beta_mega$ub = all_beta_mega$beta + all_beta_mega$std.err
  all_beta_mega$lb = all_beta_mega$beta - all_beta_mega$std.err
    
    
    
    
    
  #add the *** label for significant vs. not significant
  annotation_label_mega <- all_beta_mega
  annotation_label_mega$mark <- ifelse(annotation_label_mega$bonf <0.05,"***",
                                       ifelse(annotation_label_mega$fdr <0.1,"•", ""))

  
  
  
  my_colours = c('blue', 'grey', 'red', 'green', 'yellow', 'purple') 
    mega_analysis_plot = all_beta_mega %>% 
    ggplot(aes(x = subclass, y = beta,fill = AIBS_class_color)) + theme_minimal()+
    geom_hline(yintercept = 0) + 
    geom_bar(stat = "identity", show.legend = FALSE) + 
    scale_fill_manual(values = my_colours) + 
    facet_grid(~AIBS_class_label, scale = 'free_x', space = 'free_x') +
    geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
    ylab('LOAD (Beta)') + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
          strip.text.x = element_text(size = 15))+
    ggtitle(paste0("Mega Analysis Results for ", folder_name, " Marker List")) +
    geom_text(x = annotation_label_mega$subclass,  y = 0.4, 
              label = annotation_label_mega$mark, 
              colour = "black", size=6) +expand_limits(y = c(-0.6, 0.5))
    
  full_res$cohort <- as.character(full_res$cohort)
    

  replacement <- list(c("ROSMAP", "DLPFC"), c("BM10", "FP"), 
                   c("BM44", "IFG"),c("BM36","PHG"), c("BM22", "STG"), 
                   c("MAYO", "TCX"))
  
  for(replace in replacement){
      full_res$cohort <- gsub(replace[[1]], replace[[2]], full_res$cohort)
  }
                                      

  cohort_label  <- c("DLPFC", "FP", "IFG", "STG", "PHG", "TCX")
     
  label_mark_IT <- subset(annotation_label, celltype == "IT")
  data_text_IT <- data.frame(
        label = label_mark_IT$mark,
        cohort   = cohort_label
      )
  IT_boxplot <- full_res %>% 
        ggplot(aes(x = Diagnosis, y = ITLOADResid )) + theme_minimal()+
        #geom_violin(alpha=0.4) + 
        geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
        stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
        scale_fill_brewer(palette="Set1")+
        scale_color_brewer(palette="Set1")+
        ggtitle(paste0(' IT rCTP residuals association with LOAD for ', markers))+
        facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
        ylab('IT rCTP residuals') + 
        xlab('')  + geom_text(
        data    = data_text_IT,
        mapping = aes(x = 1, y = 3.5, label = label),
        hjust   = -0.1,
        vjust   = -1
      )+expand_limits(y = c(-8, 4))
  
  label_mark_SST <- subset(annotation_label, celltype == "SST")
  data_text_SST <- data.frame(
        label = label_mark_SST$mark,
        cohort   = cohort_label
      )
  SST_boxplot <- full_res %>% 
        ggplot(aes(x = Diagnosis, y = SSTLOADResid )) + theme_minimal()+
        #geom_violin(alpha=0.4) + 
        geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
        stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
        scale_fill_brewer(palette="Set1")+
        scale_color_brewer(palette="Set1")+
        ggtitle(paste0(' SST rCTP residuals association with LOAD for ', markers))+
        facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
        ylab('SST rCTP residuals') + 
        xlab('')  + geom_text(
        data    = data_text_SST,
        mapping = aes(x = 1, y = 4, label = label),
        hjust   = -0.1,
        vjust   = -1
      )+expand_limits(y = c(-4, 5))
  
  
     
  ## generate the final multi-panel temporal figure using cowplot's plot_grid
  bottom_plot = plot_grid(SST_boxplot, IT_boxplot,  
                      nrow = 2, ncol = 1, rel_widths = c(.6, .4), 
                      axis = 'l', align = 'v', labels = c('B', 'C'))
  full_plot = plot_grid(mega_analysis_plot, 
                            bottom_plot, nrow = 2 , 
                            rel_heights = c(1.5,2.5), labels = c('A'))
  
  saveRDS(full_plot, paste0("./figures/figure1", markers, ".rds"))
  pdf(file = paste0("./figures/figure1", markers, ".pdf"), 
      width = 10, # The width of the plot in inches
      height = 12)
  print(full_plot)
  dev.off()
  print(full_plot)
}
```