---
title: "Mega-Analysis -- FINAL snCTP"
output:
  html_document:
    df_print: paged
---
```{r echo=FALSE, cache=FALSE, comment=FALSE, warning=FALSE}
#load necessary libraries
library(magrittr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(limma)
library(edgeR)
library(rms)
library(useful)
library(stringr)
library(markerGeneProfile)
library(rlist)
library(lme4)
library(ggthemes)
library(ggsci)
library(ggpubr)
library(ggbeeswarm)
library(cowplot)
```
Now that we've loaded in the libraries lets get the marker list we need.
```{r}
subclass_CgG_MTG_common_final <-  
  readRDS('./commonMarkerLists/subclass_CgG_MTG_common_final.rds')
```
We're going to create a dataframe for mega-analysis for our marker list, which will contain all the single-nucleus relative cell-type proportion estimates (snCTPs) calculated for each subject in each of the datasets (Cain, Mathys and Zhou) with the given marker gene list, as well as their sample identifier (projid), their age at death, their sex, their pmi and their Alzheimer's diagnosis. 

```{r}
cohorts <- c("cain", "mathys", "zhou")
folder_name <-'./snCTPData/'  
    
for(cohort in cohorts){
  print(cohort)
  ctp_name <- paste0("ctp_",cohort)
  ctp_df <- read.table(paste0(folder_name, '/', ctp_name, ".txt"), sep= ",",header=TRUE)
  assign(ctp_name, ctp_df )
  cell_types <- unique(ctp_df$subclass)
  ctp_df <- ctp_df %>% spread('subclass', 'cell_type_proportion')
  ctp_df <- ctp_df %>% group_by(projid)  %>% fill(cell_types,.direction = "down") 
  ctp_df <- na.omit(ctp_df)
  ctp_df <- ctp_df %>% dplyr::select(cell_types, "projid", "msex", "LOAD", "pmi", "age_death")
  ctp_df$cohort <- cohort
  if(cohort == "cain"){
    mega_ctp <- ctp_df
  }
  else{
    mega_ctp <- rbind(mega_ctp, ctp_df)
  }
}


#save mega_ctp
saveRDS(mega_ctp, paste0(folder_name, '/', "megaCTP.rds"))
```
Now we have all the dataframes we need for the mega-analysis. Let's do it.

```{r}

folder_name <-'./snCTPData/'  
mega_ctp <- readRDS(paste0(folder_name, '/', "megaCTP.rds"))
#mega_ctp <- mega_ctp %>% dplyr::filter(LOAD !="OTHER")
mega_ctp <- mega_ctp %>% dplyr::mutate(LOAD = ifelse(LOAD == "AD", 1, 0))


covars <- c("pmi", "age_death", "msex")
colnames(mega_ctp) <- make.names(colnames(mega_ctp))
cell_types <- colnames(mega_ctp[,1:16])

pathology <- ("LOAD")
model.data <- mega_ctp

LOAD_results <- sapply(cell_types,function(celltype) {
  sapply(pathology, function(pathology) {
    print(celltype)
    form <- as.formula(paste0(celltype,"~",pathology," + ", "(1 | projid )" ,
                                " + ", "cohort" , " + ",  paste0(covars,collapse=" + "))) 
    model <- lmer(data=model.data,form)
    return(model)
  })
})

results <- sapply(cell_types,function(celltype) {
    print(celltype)
    form <- as.formula(paste0(celltype,"~" ," + ", "(1 | projid )" ," + ", "cohort" , 
                                 " +", paste0(covars,collapse=" + "))) 
    model2 <- lmer(data=model.data,form)
    return(model2)
  })

for(cell in cell_types){
  mod1Name <- paste0(cell, ".LOAD")
  mod2Name <- cell
  print(mod1Name)
  print(mod2Name)
  significance <- (anova(LOAD_results[mod1Name][[1]], results[mod2Name][[1]]))$`Pr(>Chisq)`[2]
  confInt <- confint(LOAD_results[mod1Name][[1]],level = 0.95,  oldNames=FALSE)
  upperBound <- confInt[4,2]
  lowerBound <- confInt[4,1]
  
  if(cell == cell_types[1]){
   celltype_sig <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2] ,
                                      "lowerBound" = lowerBound, "upperBound" = upperBound)
  }
  else{
    temp <- data.frame("celltype"=cell, significance, "beta" = coef(summary(LOAD_results[mod1Name][[1]]))[2,1], "std.err" = coef(summary(LOAD_results[mod1Name][[1]]))[2,2],
                       "lowerBound" = lowerBound, "upperBound" = upperBound)
    celltype_sig <- rbind(celltype_sig, temp)
  }
}

celltype_sig$fdr <- p.adjust(celltype_sig$significance, method="fdr")
celltype_sig$bonf <- p.adjust(celltype_sig$significance, method="bonferroni")
celltype_sig$SIG <- celltype_sig$fdr <0.05
celltype_sig$SIGBONF <- celltype_sig$bonf <0.05


saveRDS(celltype_sig, paste0(folder_name, "mega_results.rds"))

```
We've calculated the significance of the association between each of the cell-types and the AD diagnosis variable (LOAD) in a mega-analyis. We can now plot the results.
```{r}

folder_name <- './snCTPData/'
subclass_meta <- read.csv('./subclassMeta/subclass_meta.txt')


mega_ctp_results <- readRDS(paste0(folder_name, "mega_results.rds"))
  
  
  

mega_ctp_results <- mega_ctp_results %>% dplyr::rename(subclass = celltype)
subclass_meta$AIBS_subclass_label <- make.names(subclass_meta$AIBS_subclass_label)
all_beta_mega = merge(mega_ctp_results,
                subclass_meta, by.x = 'subclass', by.y = 'AIBS_subclass_label')
all_beta_mega$class <- as.factor(all_beta_mega$AIBS_class_label)
all_beta_mega$class <- factor(all_beta_mega$AIBS_class_label, 
                          levels = c("GABAergic", "Glutamatergic", 
                                                          "Non-neuronal"))
all_beta_mega <- arrange(all_beta_mega, class)

all_beta_mega$ub = all_beta_mega$beta + all_beta_mega$std.err
all_beta_mega$lb = all_beta_mega$beta - all_beta_mega$std.err

  
  
  
  
#add the *** label for significant vs. not significant
annotation_label_mega <- all_beta_mega
annotation_label_mega$mark <- ifelse(annotation_label_mega$significance <0.05,"*", "")
  
  
  
my_colours = c('blue', 'grey', 'red', 'green', 'yellow', 'purple') 

  mega_analysis_plot = all_beta_mega %>% 
  ggplot(aes(x = subclass, y = beta,fill = AIBS_class_color)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  scale_fill_manual(values = my_colours) + 
  facet_grid(~AIBS_class_label, scale = 'free_x', space = 'free_x') +
  geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
  ylab('LOAD (Beta)') + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
        strip.text.x = element_text(size = 15))+
  ggtitle(paste0("Mega Analysis Results for CgG and MTG Marker List")) +
  geom_text(x = annotation_label_mega$subclass,  y = 0.04, 
            label = annotation_label_mega$mark, 
            colour = "black", size=6)
 
print(mega_analysis_plot)
ggsave(paste0(folder_name, "mega_analysis_plot", ".png"))
```
Let's create case/control box plots for SST and IT rCTP residuals so we can see if there's a difference in cell type proportion changes across cohorts/brain regions.
```{r}
cohorts <- c("cain", "mathys", "zhou")
folder_name <-'./snCTPData/'  

full_res_indiv <- data.frame()
full_sig_indiv <- data.frame()

for(cohort in cohorts){
  print(cohort)
  ctp_name <- paste0("ctp_",cohort)
  ctp_df <- read.table(paste0(folder_name, '/', ctp_name, ".txt"), sep= ",",header=TRUE)
  assign(ctp_name, ctp_df)
  ctp_df <- ctp_df %>% dplyr:: mutate(subclass = make.names(subclass))
  cell_types <- unique(ctp_df$subclass)
  ctp_df <- ctp_df %>% spread('subclass', 'cell_type_proportion')
  ctp_df <- ctp_df %>% group_by(projid)  %>% fill(cell_types,.direction = "down") 
  ctp_df <- na.omit(ctp_df)
  ctp_df <- ctp_df %>% dplyr::select(cell_types, "projid", "msex", "LOAD", "pmi", "age_death")
  ctp_df$cohort <- cohort
  

  
  pathology <- "LOAD"
  covars <- c("msex", "age_death", "pmi")
  model.data <- ctp_df
  i=0
  for(celltype in cell_types){
    form <- as.formula(paste0(celltype,"~",paste0(covars,collapse=" + "))) 
    model <- lm(data=model.data,form)
    celltype_residual <- data.frame(resid(model))
    
    model.data[paste0(celltype, "LOADResid")] <- celltype_residual
    model.data$cohort <- cohort
    
    form <- as.formula(paste0(celltype,"~",pathology," + ",paste0(covars,collapse=" + "))) 
    model <- lm(data=model.data,form)
    loadp <- (coef(summary(model))[2,4])
    
    print(celltype, loadp)
    
    if(i==0){
         sig_results <- data.frame("celltype"= celltype, "cohort" = cohort, "significance" = loadp)
    }
    else{
      sig_results_curr <- data.frame("celltype"= celltype, "cohort" = cohort, "significance" = loadp)
      sig_results <- rbind(sig_results, sig_results_curr)
    }
    i= i+1
  }
  cell_type_resids <- paste(cell_types, "LOADResid", sep="")
    full_res_indiv <- rbind(full_res_indiv, model.data[, c(cell_type_resids, "projid", "cohort", "LOAD")])
  
  full_sig_indiv <- rbind(full_sig_indiv, sig_results)
  }
  


full_res_indiv$Diagnosis <- full_res_indiv$LOAD 
full_res_indiv$Diagnosis <- factor(full_res_indiv$Diagnosis, levels = c("C", "AD", "OTHER"))

saveRDS(full_res_indiv, paste0(folder_name, "full_res.rds"))
saveRDS(full_sig_indiv, paste0(folder_name, "full_sig.rds"))

```
Let's plot the case control boxplots now that we've calculated the residuals.
```{r}

cohorts <- c("cain", "mathys", "zhou")
folder_name <-'./snCTPData/'  

full_res <- readRDS(paste0(folder_name, "full_res.rds"))

full_sig <- readRDS(paste0(folder_name, "full_sig.rds"))


full_res$cohort <- factor(full_res$cohort, levels = c("cain", "mathys", "zhou"))
full_res <- arrange(full_res, cohort)

full_sig$cohort <- factor(full_sig$cohort, levels = c("cain", "mathys", "zhou"))
full_sig <- arrange(full_sig, cohort)


#add the *** label for significant vs. not significant
annotation_label <- full_sig
annotation_label$bonf <- p.adjust(annotation_label$significance, method="bonferroni")
annotation_label$fdr <- p.adjust(annotation_label$significance, method="fdr")
annotation_label$mark <- ifelse(annotation_label$significance <0.05,"*", "ns")
  
for(cell in cell_types){
  print(cell)
  label_mark <- subset(annotation_label, celltype == cell)
  data_text <- data.frame(
    label = label_mark$mark,
    cohort   = c("cain", "mathys", "zhou")
  )
  boxplots <- full_res %>% 
    ggplot(aes(x = Diagnosis, y = get(paste0(cell, "LOADResid")) )) + theme_minimal()+
    #geom_violin(alpha=0.4) + 
    geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
    stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
    scale_fill_brewer(palette="Set1")+
    scale_color_brewer(palette="Set1")+
    ggtitle(paste0(cell, ' snCTP residuals association with LOAD for CgG MTG markers'))+
    facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
    ylab(paste0(cell, ' rCTP residuals')) + 
    xlab('')  + geom_text(
    data    = data_text,
    mapping = aes(x = 1, y = 0.1, label = label),
    hjust   = -0.1,
    vjust   = -1
  )
  assign(paste0(cell, "_boxplot"), boxplots)
  print(boxplots)
  saveRDS(get(paste0(cell, "_boxplot")), 
          paste0(folder_name, "case_control_", cell, "_boxplot.rds"), compress=TRUE)

  ggsave(paste0(folder_name, "case_control_", cell, "_boxplot", ".png"))
}
  
  
  
```

```{r}

subclass_meta <- read.csv('./subclassMeta/subclass_meta.txt')


  
cohorts <- c("cain", "mathys", "zhou")
folder_name <-'./snCTPData/'  

full_res <- readRDS(paste0(folder_name, "full_res.rds"))

full_sig <- readRDS(paste0(folder_name, "full_sig.rds"))
  
cohort_name <- c("cain", "mathys", "zhou")
full_res$cohort <- factor(full_res$cohort, levels = cohort_name)
full_sig$cohort <- factor(full_sig$cohort, levels = cohort_name)
  
full_res <- arrange(full_res, cohort)
full_sig <- arrange(full_sig, cohort)
#add the *** label for significant vs. not significant
annotation_label <- full_sig
annotation_label$bonf <- p.adjust(annotation_label$significance, method="bonferroni")
annotation_label$mark <- ifelse(annotation_label$significance <0.05,"*", "ns")
  
  
  #get mega analysis plot
  mega_ctp_results <- readRDS(paste0(folder_name, "mega_results.rds"))
  

mega_ctp_results <- mega_ctp_results %>% dplyr::rename(subclass = celltype)
subclass_meta$AIBS_subclass_label <- make.names(subclass_meta$AIBS_subclass_label)
all_beta_mega = merge(mega_ctp_results,
                subclass_meta, by.x = 'subclass', by.y = 'AIBS_subclass_label')
all_beta_mega$class <- as.factor(all_beta_mega$AIBS_class_label)
all_beta_mega$class <- factor(all_beta_mega$AIBS_class_label, 
                          levels = c("GABAergic", "Glutamatergic", 
                                                          "Non-neuronal"))
all_beta_mega <- arrange(all_beta_mega, class)
all_beta_mega$ub = all_beta_mega$beta + all_beta_mega$std.err
all_beta_mega$lb = all_beta_mega$beta - all_beta_mega$std.err
  

  
all_beta_mega$subclass <- gsub(".", " ", all_beta_mega$subclass, fixed=TRUE)
all_beta_mega$subclass <- factor(all_beta_mega$subclass, 
                            levels = c("LAMP5", "PAX6", "VIP", "SST", "PVALB", "IT", 
                            "L4 IT", "L5 6 NP", "L5 ET", "L6 CT", "L5 6 IT Car3", "L6b",
                            "Astrocyte", "Endothelial", "Microglia", "Oligodendrocyte",
                            "OPC", "Pericyte", "VLMC"))
  
  
  
#add the *** label for significant vs. not significant
annotation_label_mega <- all_beta_mega
annotation_label_mega$mark <- ifelse(annotation_label_mega$significance <0.05,"*", "")
  
  
  
my_colours = c('blue', 'grey', 'red', 'green', 'yellow', 'purple') 
mega_analysis_plot = all_beta_mega %>% 
ggplot(aes(x = subclass, y = beta,fill = AIBS_class_color)) + theme_minimal()+
geom_hline(yintercept = 0) + 
geom_bar(stat = "identity", show.legend = FALSE) + 
scale_fill_manual(values = my_colours) + 
facet_grid(~AIBS_class_label, scale = 'free_x', space = 'free_x') +
geom_errorbar(aes(ymin = lb, ymax = ub), width = .33) + 
ylab('LOAD (Beta)') + 
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 15),
      strip.text.x = element_text(size = 15))+
ggtitle(paste0("Mega Analysis Results for CgG and MTG Marker List")) +
geom_text(x = annotation_label_mega$subclass,  y = 0.04, 
          label = annotation_label_mega$mark, 
          colour = "black", size=6)
     
label_mark_IT <- subset(annotation_label, celltype == "IT")
data_text_IT <- data.frame(
      label = label_mark_IT$mark,
      cohort   = c("cain", "mathys", "zhou")
    )
IT_boxplot <- full_res %>% 
      ggplot(aes(x = Diagnosis, y = ITLOADResid )) + theme_minimal()+
      #geom_violin(alpha=0.4) + 
      geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
      stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
      scale_fill_brewer(palette="Set1")+
      scale_color_brewer(palette="Set1")+
      ggtitle("IT rCTP residuals association with LOAD for CgG and MTG Marker List")+
      facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
      ylab('IT rCTP residuals') + 
      xlab('')  + geom_text(
      data    = data_text_IT,
      mapping = aes(x = 1, y = 0.1, label = label),
      hjust   = -0.1,
      vjust   = -1
    )
  
label_mark_SST <- subset(annotation_label, celltype == "SST")
data_text_SST <- data.frame(
      label = label_mark_SST$mark,
      cohort   = c("cain", "mathys", "zhou")
    )
SST_boxplot <- full_res %>% 
      ggplot(aes(x = Diagnosis, y = SSTLOADResid )) + theme_minimal()+
      #geom_violin(alpha=0.4) + 
      geom_quasirandom(size=0.6,shape=19,aes(col=Diagnosis),alpha=0.8) +
      stat_summary(fun=mean,geom="point",aes(fill=Diagnosis),size=4,shape=23,col="black")+
      scale_fill_brewer(palette="Set1")+
      scale_color_brewer(palette="Set1")+
      ggtitle("SST rCTP residuals association with LOAD CgG and MTG Marker List")+
      facet_wrap(~cohort, scales = 'free_x',nrow=1) + 
      ylab('SST rCTP residuals') + 
      xlab('')  + geom_text(
      data    = data_text_SST,
      mapping = aes(x = 1, y = 0.03, label = label),
      hjust   = -0.1,
      vjust   = -1
    )
  
  
     
## generate the final multi-panel temporal figure using cowplot's plot_grid
bottom_plot = plot_grid(SST_boxplot, 
                    IT_boxplot,  
                    nrow = 2, ncol = 1, rel_widths = c(.6, .4), 
                    axis = 'l', align = 'v', labels = c('B', 'C'))
full_plot = plot_grid(mega_analysis_plot, 
                          bottom_plot, nrow = 2 , 
                          rel_heights = c(1.5,2.5), labels = c('A'))
markers <- "subclass_CgG_MTG"
saveRDS(full_plot, paste0("./figures/figure2", markers, ".rds"))
pdf(file = paste0("./figures/figure2", markers, ".pdf"), 
    width = 10, # The width of the plot in inches
    height = 12)
print(full_plot)
dev.off()
print(full_plot)
```